%%%%%%%%%%%%%%%%%%%%%%%% ACK %%%%%%%%%%%%%%%%%%%

\pagenumbering{roman}
\setcounter{page}{25}


%#############################################################
\chapter*{Acknowledgments}
%#############################################################

\addcontentsline{toc}{chapter}{\protect\numberline{}{Acknowledgments}}
\thispagestyle{plain}
\markboth{ACKNOWLEDGMENTS}{ACKNOWLEDGMENTS}

\noindent
Thank you to the many students who have taken our Sustainability Challenges course 
over the years and engaged deeply with course content.
Thank you to our colleague Julie Wildschut for teaching Sustainability Challenges in Fall 2021
and prototyping this text throughout the semester.

Thank you to student researchers Joshua Broekhuisen, Anjana Sainju, 
Henos Tadesse, and Larisa Tomeci
for help with sourcing data for the figures in the text. 

Serveral reviewers provided feedback on early versions to the book---thank you!
James VanAntwerp provided feedback on an early version of Part~\ref{part:IPARX}.
Margaret VanAntwerp provided extensive feedback as a student in the
Sustainability Challenges course in Fall 2021.
% Other reviewers were Julie Wildschut and Glen VanAntwerp. Becky Haney, James Skillen, Uko Zylstra.
Other reviewers were Paul Brockway, Nathan Grawe, Zeke Marshall,
Gregor Semieniuk, T\^{a}nia Sousa, and Uko Zylstra.

% Jeff Jewett, Shannon Savage, Leonides Murembya, Steve McMullen, Nathan Grawe
% Jennifer VanAntwerp,  Glen VanAntwerp, ...
% Derek Schuurman, Sa BA, Tom Ackerman

Thank you to the Calvin University Board of Trustees and 
to Joel and Linda Zylstra
for funding to create and revise the text.
Thanks also to Tim Koning of LHM Tank for providing funding for biofuels research.



\vspace*{2pc}
\noindent\AUTHORS\\
\noindent January 2022