%%%%%%%%%%%%%%%%%%%%%%%% PREFACE %%%%%%%%%%%%%%%%%%%

\pagenumbering{roman}
\setcounter{page}{21}

{
  \chapter*{Preface}
  \addcontentsline{toc}{chapter}{\protect\numberline{}{Preface}}
  \thispagestyle{plain}
  \markboth{PREFACE}{PREFACE}
  
  \noindent
  
  
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \section*{Who should use this book} 
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    
    This book is for anyone who is interested in sustainability.
  It introduces the many challenges of sustainability
  and provides a framework for thinking about sustainability.
  The reader is not assumed to have prior background knowledge,
  apart from basic numeracy and the ability to comprehend graphical information.
  This book is not intended to be comprehensive in its coverage of all aspects of 
  sustainability.
  Neither does it suggest solutions to many sustainability problems. 
  It should be easy to read but hard to digest.
  
  
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \section*{How to use this book} 
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    
    This book was written to support a university seminar class %one semester-hour 
  on sustainability. 
  The primary focus of class time is in-class discussion, so 
  the end-of-chapter discussion questions should provide significant value. 
  Therefore, this book could be used in other settings that prioritorize discussion,
  like a book club or small group.
  Individual chapters, or sets of chapters, can easily be
  used to support a wide variety of class structures, types, and levels.
  Students should read the text on their own, out of class, to get 
  basic information related to  the challenges of sustainability. 
  We encourage instructors to use a basic and straightforward quiz on the reading 
  through your course management platform to hold students accountable for the 
  reading. 
  Students should review discussion questions ahead of each class and 
  prepare for discussion.
  
  
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \section*{What you should put into this book} 
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    
    It is often said that ``you only get out what you put in'' to an activity.
  We encourage readers to be critical.
  What other sustainability concerns do you have?
    What other ideas can you suggest to become more sustainable?
    
    As you read, we encourage you to keep the table of key terms in Table~\ref{tab:key_terms} at hand.
  
  
  {
    
    % This next command adjusts table numbering for this table 
    % and its reference only.
    % It strips out the chapter number, 
    % thereby avoiding "Table 0.1".
    % Note, this works for this table only, because 
    % the command is sandwiched between two curly braces.
    \renewcommand{\thetable}{\arabic{table}}
    
    \begin{table}[h!]
    \caption[Key terms]{Key terms.}
    \centering
    \begin{tabular}{r l}
    \toprule
    Key term              & Meaning  \\ 
    \midrule  
    The economy           & Society's metabolism. \vspace{4mm} \\
%
\multirow{2}{*}{Effectiveness}  & A general term that encompasses efficacy, \\
                                & efficiency, EROI, and productivity. \vspace{4mm} \\
%
\multirow{2}{*}{Efficacy}   & The ratio $\frac{\text{useful output}}{\text{input}}$, with different units \\
                            & in numerator and denominator. \vspace{4mm} \\
%
Efficiency            & A unitless ratio $\frac{\text{useful output}}{\text{input}}$, often for energy. \vspace{4mm} \\
%
\multirow{3}{*}{Energy service}   & The purpose of energy consumption, \\
                                  & for example transportation, illumination, \\
                                  & or space heating. \vspace{4mm} \\
%
\multirow{3}{*}{EROI} & Energy return on investment, a key metric for \\
                      & energy production systems, defined as $\frac{\text{energy produced}}{\text{energy input}}$, \\
                      & where ``energy input'' does not include feedstock. \vspace{4mm} \\
%
\multirow{2}{*}{Externality}   & A cost or benefit borne by a third party \\ 
                               & to an economic transaction. \vspace{4mm} \\
%
\multirow{2}{*}{Intensity}  & The ratio $\frac{\text{input}}{\text{output}}$, for example
                              $\mft{R} = \frac{\text{resource consumption}}{\text{GDP}}$, \\
                            & with different units in numerator and denominator. \vspace{4mm} \\
%
\multirow{3}{*}{Productivity} & The ratio $\frac{\text{economic output (\$)}}{\text{production input}}$, a type of efficacy, \\
                              & often with different units in numerator~(\$) and  \\
                              & denominator (capital, labor, energy, or materials). \vspace{4mm} \\
%
\multirow{2}{*}{Resources}  & Material extracted from the natural environment \\
                            & for purposes of human consumption. \vspace{4mm} \\
%
\multirow{2}{*}{Strong sustainability} & Using resources more slowly than \\
                                       & natural regeneration rates. \vspace{4mm} \\
%
Weak sustainability         & Using resources so they last 50~years. \\
\bottomrule
\end{tabular}
\label{tab:key_terms}
\end{table}

} % End of changing \thetable command.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{What you should get out of this book} 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%t

This book contains four main contributions.

\begin{itemize}

  \item Basic facts, figures, and information related to sustainability. 
        We don't assume you start with any prior knowledge, but by the time you finish 
    reading this book you will know basic facts like carbon dioxide 
    concentration in the atmosphere. 
    Furthermore, you should gain a sense of scale related to sustainability 
    information. 
    That is, which things are big and important and 
    which are smaller and less relavant. 
    To this end, information is presented graphically (as much as possible) 
    to show relative magnitudes and trends.
    
    \item An explanatory framework for thinking about sustainability,
    primarily in chapters \ref{chap:introduction}--\ref{chap:impact_intensity}.
    
    \item A wealth of end-of-chapter discussion questions that 
    address moral, ethical, philosophical,
    and practical aspects of sustainability. 
    Most questions don't have objective answers. 
        Instead, they illustrate the importance
        of worldview, values, preferences, and \emph{a priori} assumptions 
        for each person's approach to sustainability. 
    
    \item Inspiration for future sustainability-related studies.
    The end-of-chapter projects range in scope from a long
    homework problem to a graduate thesis. 
    Instructors can use one---or perhaps more---of the projects for
    semester-long or capstone assignmets,
    with optional in-class presentations.
    
    \end{itemize}
    
    When you finish Chapter~\ref{chap:personal_action}, you should see
    that sustainability entails important, urgent, and difficult challenges; 
    that sustainability transitions are needed; and 
    that you can play a role in helping the world become more sustainable
    through a variety of individual and corporate actions.
    
    
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
      \section*{Text, organization, and schedule} 
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%t
    
    At one chapter a week, the twelve chapters in this book don't quite fill a typical 
university semester. 
Of course, you don't have to be in a university context to use this book. 
    A book club or small group are also appropriate audiences---sustainability 
    is important for everyone. 
    However, if you are in a university context, here are a few ways that you could fill the 
    remaining weeks of your semester. 
    Use class time for project presentations. 
    Perhaps groups of students could each be assigned different project questions and, 
    at the end of the semester, each could share their results.
    Alternatively, one or more class sessions could be devoted to a topic or topics 
    of the instructor's choice, like biofuels, grid-scale storage, 
or media coverage of sustainability. 
At Calvin University, our class devotes a week to worldview implications for 
sustainability. 
You could use all or part(s) of \emph{Beyond Stewardship: New approaches to 
creation care}~\citep{Warners:2019aa}
or \emph{Current disciplines and worldviews are insufficient
to address sustainability challenges}~\cite{VanH2019} as a starting point for discussion. 
On the other hand, instructors don't need to use all of the book.
%IEEE Xplore makes it easy to assign 
Either a single chapter or a group of chapters may be used as supplementary 
material for many different types of classes.
    Now, on to the good stuff!
      
      \vspace*{2pc}
    \noindent\AUTHORS\\
    \noindent January 2022
  }
  
  
  \bibliographystyle{unsrtnat}
  \bibliography{MCBook2021}
  
  % \input{fm.bbl}
  
  
  \cleardoublepage